---
title: "simpleRegress"
author: "Melanie Espinosa"
date: "Spring 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 
```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
```{r}
Ourweight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(Ourweight)
summary.lm(Ourweight)
```
based on what we have seen the neck is better then the weight 
Weight: RSE: 5.432, ARs: 0.7423
Neck: RSE: 5.376 Ars: 0.7476

```{r}
Ourchest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(Ourchest)
summary.lm(Ourchest)
```
The order from worst to best would be weight, neck, chest. 

Weight: RSE: 5.342 ARs:0.7423
Neck: RSE: 5.376 ARs: 0.7476
Chest: RSE: 4.939 ARS: 0.7869

```{r}
ourheadwth <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(ourheadwth)
summary.lm(ourheadwth)
```

the order from worst to best be HEAFWTH, WEIGHT,NECK and finally CHEST.
Headwth RSE: 7.32 ARs: 0.5321
Weight: RSE: 5.342 ARs:0.7423
Neck: RSE: 5.376 ARs: 0.7476
Chest: RSE: 4.939 ARS: 0.7869


```{r}
Ourheadlen <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(Ourheadlen)
summary.lm(Ourheadlen)
```

the worse to best starting with Headwth, Weight, Neck, Chest and finally Headlen
Headwth RSE: 7.32 ARs: 0.5321
Weight: RSE: 5.342 ARs:0.7423
Neck: RSE: 5.376 ARs: 0.7476
Chest: RSE: 4.939 ARS: 0.7869
Headlen: RSE: 4.235 ARs: 0.8434









